---
title: "README"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

  <!-- badges: start -->
  [![tic](https://github.com/IhsanKhaliq/ascotraceR/workflows/tic/badge.svg?branch=master)](https://github.com/IhsanKhaliq/ascotraceR/actions)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
<!-- badges: end -->

# ascotraceR: A weather driven model to simulate the spread of Ascochyta blight in chickpea over a growing season 

The goal of of _ascotraceR_ is to develop a weather driven model to simulate the spread of Ascochyta blight disease in a chickpea field over a growing season.

This model is adapted from a model developed by [(Diggle _et al._ 2002)](https://doi.org/10.1094/PHYTO.2002.92.10.1110) for simulating the spread of anthracnose in a lupin field.
The model is run using local weather data, and simulates both host and pathogen related activities.
The growth of chickpea is described in terms of development of growing points, and disease progress is measured by quantifying the loss of healthy growing points.
As the model runs, it keeps a continuous track of non-infected, latent, infected and sporulating growing points (lesions).
The ascotraceR's minimum input requirements are location specific weather data and a list of input variables.

## Quick start

_ascotraceR_ is not yet on CRAN.
You may install it from GitHub this way.

```r
if (!require("remotes"))
  install.packages("remotes")
remotes::install_github("IhsanKhaliq/ascotraceR",
                        build_vignettes = TRUE
)
```

Once installed you can simulate disease spread in a chickpea paddock.

Load the library.

```{r, load-libs, message=FALSE}
library("ascotraceR")
library("data.table")
library("lubridate")
```

Import the weather data

```{r load-weather}
# weather data
Billa_Billa <- fread(
  system.file(
    "extdata",
    "2020_Billa_Billa_weather_data_ozforecast.csv",
    package = "ascotraceR"
  )
)

# format time column
Billa_Billa[, local_time := dmy_hm(local_time)]

# specify the station coordinates of the Billa Billa weather station
Billa_Billa[, c("lat", "lon") := .(-28.1011505, 150.3307084)]

head(Billa_Billa)
```

### Wrangle weather data 

A function, `format_weather()`, is provided to convert raw weather data into the format appropriate for the model.
It is mandatory to use this function to ensure weather data is properly formatted before running the model.

```{r format-weather}
Billa_Billa <- format_weather(
  x = Billa_Billa,
  POSIXct_time = "local_time",
  temp = "mean_daily_temp",
  ws = "ws",
  wd_sd = "wd_sd",
  rain = "rain_mm",
  wd = "wd",
  station = "location",
  time_zone = "Australia/Brisbane",
  lon = "lon",
  lat = "lat"
)
```

### Simulate Ascochyta blight spread

A function, `trace_asco()`, is provided to simulate the spread of Ascochyta blight in a chickpea field over a growing season.

```{r trace-asco}
# Predict Ascochyta blight spread for the year 2020 at Billa Billa
traced <- trace_asco(
  weather = Billa_Billa,
  paddock_length = 20,
  paddock_width = 20,
  initial_infection = "2020-07-17",
  sowing_date = "2020-06-04",
  harvest_date = "2020-10-27",
  time_zone = "Australia/Brisbane",
  seeding_rate = 40,
  gp_rr = 0.0065,
  spores_per_gp_per_wet_hour = 0.6,
  latent_period_cdd = 150,
  primary_inoculum_intensity = 100,
  primary_infection_foci = "centre"
)
```

### Summarise the output

You can easily get summary statistics for the whole paddock over the simulated season and area under the disease progress curve, AUDPC, using `summarise_trace()`.

```{r summarise}
summarise_trace(traced)
```

## Reference

> Diggle AJ, Salam MU, Thomas GJ, Yang H, O'connell M, Sweetingham M, 2002. AnthracnoseTracer: a spatiotemporal model for simulating the spread of anthracnose in a lupin field. Phytopathology 92, 1110-21.
